State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-04-14 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
adding to resutls

Considering the spectral representation of the dynamic field at time $t$ denoted by $V_t(\boldsymbol{\nu})$, multi-dimensional sampling and reconstruction theory dictates that for and accurate representation, $V_t(\boldsymbol{\nu})$ need to be spatially bandlimited. Such a condition is usually satisfied due to the bandlimiting effect of most practical kernels and the spatially colored covariance functions \todo{~cite\{some eeg literature here\}}. In addition, the Gaussian pick-up of the sensors provides a low-pass filtering effect, further limiting the effective bandwidth of the field. Thus the spectral response $V_t(\boldsymbol{\nu})$ is usually band-limited with $V_t(\boldsymbol{\nu}) \approx 0 ~ \forall \boldsymbol{\nu} > \boldsymbol{\nu}_c$, where $\boldsymbol{\nu}_c$ is a cutoff frequency. Note that for the 3-dimensional, homogeneous and isotopic fields being considered $\boldsymbol{\nu}_c = [\nu_c ~ \nu_c ~ \nu_c]^\top$. Given such a band-limited field and based on Shannon's sampling theorem, the distance between adjacent sensors, $\Delta_s$, needs to satisfy
\begin{equation}
\Delta_s = \frac{1}{2\rho\nu_c},
\end{equation}
where $\rho \in \mathbb{R} \ge 1$ is an oversampling parameter. Satisfying this condition is required for the hidden dynamic field $v_t(\mathbf{r})$ to be precisely inferred from the sampled observations, $\mathbf{y}(t)$. If the condition is not satisfied, spatial aliasing will occur.

@@

The approximation of the state density is made by propagating the so called sigma points through the state-space model. The sigma points describe the mean and covariance of the state estimate. The number of sigma points needed is $2L+1$ where $ L$ is the dimension of the state space. The  columns of the sigma point matrix and their corresponding weights used in the UKF can be calculated as
\begin{equation}
 \mathcal X_{0}=\bar x
\end{equation}
\begin{equation}
 \mathcal X_{i}=\bar x+(\sqrt{( L + \lambda)\mathbf P_x})_i \quad i=1, \dots,  L
\end{equation}
\begin{equation}
 \mathcal X_{i}=\bar x-(\sqrt{( L + \lambda)\mathbf P_x})_{i- L} \quad i= L+1, \dots, 2 L
\end{equation}
\begin{equation}
 \mathbf W_0^{(m)}=\frac{\lambda}{ L+\lambda}
\end{equation}
\begin{equation}
 \mathbf W_0^{(c)}=\frac{\lambda}{ L+\lambda}+(1-\alpha^2+\beta)
\end{equation}
\begin{equation}
 \mathbf W_i^{(m)}=\mathbf W_i^{(c)}=\frac{1}{2( L+\lambda)} \quad i=1, \dots, 2L
\end{equation}
where the scaling parameter $\lambda$ is defined as
\begin{equation}
\lambda=\alpha^2( L+\kappa)- L 
\end{equation}
and parameters $\alpha$, $\beta$ and $\kappa$ are used to tune the filtering process \cite{Haykin2001}. It should be noted that $\bar x$ is $ \hat{\mathbf x}_t^{f} $ and $ \hat{\mathbf x}_t^{b} $ with their corresponding covariance matrices in filtering and smoothing iterations respectively.

@@

\end{algorithmic}
\end{algorithm}
\section{Results}
The neural field model was used to the generate data for state and parameter estimation. All parameters for the model are given in Table~\ref{}. An example of the simulated field can be seen in Figure~\ref{}~A. The Figure demonstrates the experimental design where the sensors were placed in the centre of the field. This design allowed for uncertainty around the edges of the sensor array, which would be typical from physical data, and alleviated problems associated with boundary conditions. Figure~\ref{}~B. shows example data from the sensors of the model, and Figure~\ref{}~C. shows the power spectral density of on of the simulated time-series. The PSD shows the typical $1/f$ characteristics of intracranial EEG.
% \begin{figure}
% 	\centering
% 		\includegraphics[scale=1]{file.eps}
% 	\caption{Example of experimental design and model output. \textbf{A}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the width ($1/2$ height of Gaussian) are shown by the circles. \textbf{B} Example of data generated by the model through the observations. \textbf{C}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.}
% 	\label{fig:ExampleFieldAndSensorsDataAndPSD}
% \end{figure}

\subsection{Spatial Frequency Analysis}

>>>>

Considering the spectral representation of the dynamic field at time $t$ denoted by $V_t(\boldsymbol{\nu})$, multi-dimensional sampling and reconstruction theory dictates that for and accurate representation, $V_t(\boldsymbol{\nu})$ need to be spatially band-limited. Such a condition is usually satisfied due to the band-limiting effect of most practical kernels and the spatially coloured covariance functions \todo{~cite\{some EEG literature here\}}. In addition, the Gaussian pick-up of the sensors provides a low-pass filtering effect, further limiting the effective bandwidth of the field. Thus the spectral response $V_t(\boldsymbol{\nu})$ is usually band-limited with $V_t(\boldsymbol{\nu}) \approx 0 ~ \forall \boldsymbol{\nu} > \boldsymbol{\nu}_c$, where $\boldsymbol{\nu}_c$ is a cutoff frequency. Note that for the 3-dimensional, homogeneous and isotopic fields being considered $\boldsymbol{\nu}_c = [\nu_c ~ \nu_c ~ \nu_c]^\top$. Given such a band-limited field and based on Shannon's sampling theorem, the distance between adjacent sensors, $\Delta_s$, needs to satisfy
\begin{equation}
\Delta_s = \frac{1}{2\rho\nu_c},
\end{equation}
where $\rho \in \mathbb{R} \ge 1$ is an oversampling parameter. Satisfying this condition is required for the hidden dynamic field $v_t(\mathbf{r})$ to be precisely inferred from the sampled observations, $\mathbf{y}(t)$. If the condition is not satisfied, spatial aliasing will occur.

@@

The approximation of the state density is made by propagating the so called sigma points through the state-space model. The sigma points describe the mean and covariance of the state estimate. The number of sigma points needed is $2L+1$ where $L$ is the dimension of the state space. The columns of the sigma point matrix and their corresponding weights used in the UKF can be calculated as
\begin{equation}
 \mathcal X_{0}=\bar x
\end{equation}
\begin{equation}
 \mathcal X_{i}=\bar x+(\sqrt{( L + \lambda)\mathbf P_x})_i \quad i=1, \dots,  L
\end{equation}
\begin{equation}
 \mathcal X_{i}=\bar x-(\sqrt{( L + \lambda)\mathbf P_x})_{i- L} \quad i= L+1, \dots, 2 L
\end{equation}
\begin{equation}
 \mathbf W_0^{(m)}=\frac{\lambda}{ L+\lambda}
\end{equation}
\begin{equation}
 \mathbf W_0^{(c)}=\frac{\lambda}{ L+\lambda}+(1-\alpha^2+\beta)
\end{equation}
\begin{equation}
 \mathbf W_i^{(m)}=\mathbf W_i^{(c)}=\frac{1}{2( L+\lambda)} \quad i=1, \dots, 2L
\end{equation}
where the scaling parameter $\lambda$ is defined as
\begin{equation}
\lambda=\alpha^2( L+\kappa)- L 
\end{equation}
and parameters $\alpha$, $\beta$ and $\kappa$ are used to tune the filtering process \cite{Haykin2001}. It should be noted that $\bar x$ is $ \hat{\mathbf x}_t^{f} $ and $ \hat{\mathbf x}_t^{b} $ with their corresponding covariance matrices in filtering and smoothing iterations respectively.

@@

\end{algorithmic}
\end{algorithm}
\section{Results}
The neural field model was used to the generate data for state and parameter estimation. All parameters for the model are given in table~\ref{}. An example of the simulated field can be seen in figure~\ref{}~A. The Figure demonstrates the experimental design where the sensors were placed in the centre of the field. This design allowed for uncertainty around the edges of the sensor array, which would be typical from physical data, and alleviated problems associated with boundary conditions. Figure~\ref{}~B. shows example data from the sensors of the model, and figure~\ref{}~C. shows the power spectral density of on of the simulated time-series. The PSD shows the typical $1/f$ characteristics of intracranial EEG.
% \begin{figure}
% 	\centering
% 		\includegraphics[scale=1]{file.eps}
% 	\caption{Example of experimental design and model output. \textbf{A}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the width ($1/2$ height of Gaussian) are shown by the circles. \textbf{B} Example of data generated by the model through the observations. \textbf{C}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.}
% 	\label{fig:ExampleFieldAndSensorsDataAndPSD}
% \end{figure}

\subsection{Spatial Frequency Analysis}
The spatial frequency of the observations is shown in figure~\ref{}. The cut-off frequency was taken to correspond to half the peak amplitude in dB and the oversampling parameter $\rho=3$. From equation~\ref{}, this gives a minimum distance between basis functions of \todo{??} and width of \todo{??} to satisfy Shannon's sampling criterion. Following this, the distance between basis function, $\Delta_b$, was set to \todo{??} and the basis function width, $\sigma$ was set to \todo{??}.
